---
title: "Final Project Documentation"
output:
  html_document: default
  word_document: default
  pdf_document: default
author: 'Brandon Sucrese'
date: '2023-04-08'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Imports, include=FALSE}
library('tswge')
library('tidyr')
library('dplyr')
library('nnfor')
library('vars')
library('RColorBrewer')
```


######################
## Data Preparation ##
######################

#### Data Importing & Setup ####
```{r Data Fixing}
data = read.csv("https://raw.githubusercontent.com/SucrazyGit/TS-Final-Project/main/Turbine_Data.csv")
data.new = data[seq(1, nrow(data), 6), ] ## Getting Data Collected Every Hour
wind = data.new[19405:19704,] ## Getting only relevant data

wsTrain = wind[1:276,]
wsTest = wind[277:300,]
wlTrain = wind[1:228,]
wlTest = wind[229:300,]
```

#########
## EDA ##
#########

#### Power Data ####
```{r Power Data Plot}
plotts.wge(wind$ActivePower, ylab = "Power", main = "Hourly Wind Turbine Power Data")
```

#### Wind Speed Data ####
```{r  Wind Speed Data Plot}
plotts.wge(wind$WindSpeed, ylab = "Wind Speed", main = "Hourly Wind Turbine Wind Speed Data")
```

#### Sample Plots ####
```{r Sample Plots, results='hide', fig.keep='all'}
plotts.sample.wge(wind$ActivePower)
```

#### ACF Plots ####
```{r ACF Plots, results='hide', fig.keep='all'}
acf(wind$ActivePower)
```

#### Spectral Density Plots ####
```{r Spectral Density Plots, results='hide', fig.keep='all'}
parzen.wge(wind$ActivePower)
```

#########################
## Wind Speed Analysis ##
#########################

#### Short Term Wind Speed Predictions ####
```{r Short Term Wind Speed Forecast}
ws.ws.mlp.fit = mlp(ts(wsTrain$WindSpeed, frequency = 24), reps = 50, comb = "median")
ws.ws.mlp.fore = forecast(ws.ws.mlp.fit, h = 24)
plotts.wge(ws.ws.mlp.fore$mean)
```

#### Long Term Wind Speed Predictions ####
```{r Long Term Wind Speed Forecast}
wl.ws.mlp.fit = mlp(ts(wlTrain$WindSpeed, frequency = 24), reps = 50, comb = "median")
wl.ws.mlp.fore = forecast(wl.ws.mlp.fit, h = 72)
plotts.wge(wl.ws.mlp.fore$mean)
```

######################################
## Active Power Univariate Analysis ##
######################################

#### Root Calculations ####
```{r Root Calculations}
factor.wge(c(rep(0,23),1)) 
factor.wge(wind$ActivePower)
wind.s24 = artrans.wge(wind$ActivePower, c(rep(0,23), 1))
```

#### Seasonality Plots ####
```{r Sample Plot, results='hide', fig.keep='all'}
plotts.sample.wge(wind.s24)
```
```{r ACF Plot, results='hide', fig.keep='all'}
acf(wind.s24,lag = 30)
```

### Small Term Predictions ###

#### ARMA(0,0,0) s=24 Model ####
```{r  ARMA(0,0,0) Short Term Model Building}
wsTrain.s24 = artrans.wge(wsTrain$ActivePower, c(rep(0,23), 1)) ## Removing Seasonality of 24
wsTrain.s24.aic = aic.wge(wsTrain.s24, type = 'bic')  ## AIC and BIC pick AR(1)
wsTrain.s24.aic
```

#### ARMA(0,0,0) s=24 Forecast ####
```{r ARMA(0,0,0) Short Term Forecast, results='hide', fig.show = 'hide'}
wsTrain.arimaOne.fore = fore.arima.wge(wsTrain$ActivePower, s = 24, n.ahead = 24)
```

#### ARMA(0,0,0) s=24 Metrics ####
```{r ARMA(0,0,0) Short Term Metrics, results='hide', fig.show = 'hide'}
ws.arimaOne.mse = mean((wsTest$ActivePower - wsTrain.arimaOne.fore$f) ^2)
ws.arimaOne.mse
ws.rollingOne.rmse = roll.win.rmse.wge(wind$ActivePower, horizon = 24, s=24)
```
```{r ARMA(0,0,0) Short Term Print Metrics}
paste0("ARMA(0,0,0) Short Term Predictions ASE: ", ws.arimaOne.mse)
paste0("ARMA(0,0,0) Short Term Rolling Predictions RMSE: ", ws.rollingOne.rmse$rwRMSE)
```

#### ARMA(0,0,0) s=24 Plots ####
```{r ARMA(0,0,0) Short Term Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "24 Hour Power Forecast")
lines(seq(277,300,1), wsTrain.arimaOne.fore$f, type = "l", col = "red")
```


#### ARMA(1,0,0) s=24 Model ####
```{r ARMA(1,0,0) Short Term Model Building}
wsTrain.s24 = artrans.wge(wsTrain$ActivePower, c(rep(0,23), 1)) ## Removing Seasonality of 24
wsTrain.s24.aic = aic.wge(wsTrain.s24, type = 'bic')  ## AIC and BIC pick AR(1)
wsTrain.s24.aic
```

#### ARMA(1,0,0) s=24 Forecast ####
```{r ARMA(1,0,0) Short Term Forecast, results='hide', fig.show = 'hide'}
wsTrain.arimaTwo.fore = fore.arima.wge(wsTrain$ActivePower, phi = wsTrain.s24.aic$p, s = 24, n.ahead = 24)
```

#### ARMA(1,0,0) s=24 Metrics ####
```{r ARMA(1,0,0) Short Term Metrics, results='hide', fig.show = 'hide'}
ws.arimaTwo.mse = mean((wsTest$ActivePower - wsTrain.arimaTwo.fore$f) ^2)
ws.arimaTwo.mse
ws.rollingTwo.rmse = roll.win.rmse.wge(wind$ActivePower, p = wsTrain.s24.aic$p, horizon = 24, s=24)
```
```{r ARMA(1,0,0) Short Term Print Metrics}
paste0("ARMA(1,0,0) Short Term Predictions ASE: ", ws.arimaTwo.mse)
paste0("ARMA(1,0,0) Short Term Rolling Predictions RMSE: ", ws.rollingTwo.rmse$rwRMSE)
```

#### ARMA(1,0,0) s=24 Plots ####
```{r ARMA(1,0,0) Short Term Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "24 Hour Power Forecast")
lines(seq(277,300,1), wsTrain.arimaTwo.fore$f, type = "l", col = "red")
```

### Long Term Predictions ###

#### ARMA(0,0,0) s=24 Model ####
```{r ARMA(0,0,0) Long Term Model Building}
wlTrain.s24 = artrans.wge(wlTrain$ActivePower, c(rep(0,23), 1)) ## Removing Seasonality of 24
wlTrain.s24.aic = aic.wge(wlTrain.s24, type = 'bic')  ## AIC and BIC pick AR(1)
wlTrain.s24.aic
```

#### ARMA(0,0,0) s=24 Forecast ####
```{r ARMA(0,0,0) Long Term Forecast, results='hide', fig.show = 'hide'}
wlTrain.arimaOne.fore = fore.arima.wge(wlTrain$ActivePower, s = 24, n.ahead = 72)
```

#### ARMA(0,0,0) s=24 Metrics ####
```{r ARMA(0,0,0) Long Term Metrics, results='hide', fig.show = 'hide'}
wl.arimaOne.mse = mean((wlTest$ActivePower - wlTrain.arimaOne.fore$f) ^2)
wl.arimaOne.mse
wl.rollingOne.rmse = roll.win.rmse.wge(wind$ActivePower, horizon = 72, s=24)
```
```{r ARMA(0,0,0) Long Term Print Metrics}
paste0("ARMA(0,0,0) Long Term Predictions ASE: ", ws.arimaOne.mse)
paste0("ARMA(0,0,0) Long Term Rolling Predictions RMSE: ", ws.rollingOne.rmse$rwRMSE)
```

#### ARMA(0,0,0) s=24 Plots ####
```{r ARMA(0,0,0) Long Term Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "72 Hour Power Forecast")
lines(seq(229,300,1), wlTrain.arimaOne.fore$f, type = "l", col = "red")
```

#### ARMA(1,0,0) s=24 Model ####
```{r ARMA(1,0,0) Long Term Model Building}
wlTrain.s24 = artrans.wge(wlTrain$ActivePower, c(rep(0,23), 1)) ## Removing Seasonality of 24
wlTrain.s24.aic = aic.wge(wlTrain.s24, type = 'bic')  ## AIC and BIC pick AR(1)
wlTrain.s24.aic
```

#### ARMA(1,0,0) s=24 Forecast ####
```{r ARMA(1,0,0) Long Term Forecast, results='hide', fig.show = 'hide'}
wlTrain.arimaTwo.fore = fore.arima.wge(wlTrain$ActivePower, phi = wlTrain.s24.aic$p, s = 24, n.ahead = 72)
```

#### ARMA(1,0,0) s=24 Metrics ####
```{r ARMA(1,0,0) Long Term Metrics, results='hide', fig.show = 'hide'}
wl.arimaTwo.mse = mean((wlTest$ActivePower - wlTrain.arimaTwo.fore$f) ^2)
wl.arimaTwo.mse
wl.rollingTwo.rmse = roll.win.rmse.wge(wind$ActivePower, phi = wlTrain.s24.aic$p, horizon = 72, s=24)
```
```{r ARMA(1,0,0) Long Term Print Metrics}
paste0("ARMA(0,0,0) Long Term Predictions ASE: ", ws.arimaOne.mse)
paste0("ARMA(0,0,0) Long Term Rolling Predictions RMSE: ", ws.rollingOne.rmse$rwRMSE)
```

#### ARMA(1,0,0) s=12 Plots ####
```{r ARMA(1,0,0) Long Term Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "72 Hour Power Forecast")
lines(seq(229,300,1), wlTrain.arimaTwo.fore$f, type = "l", col = "red")
```

###############################
## Active Power VAR Analysis ##
###############################

### Small Term Predictions ###

#### VAR Model Setup ####
```{r VAR Short Term Model Setup}
wsTrain.ap = wsTrain$ActivePower
wsTrain.ws = wsTrain$WindSpeed
ws.train.df = c(wsTrain.ap,wsTrain.ws)
```

#### VAR Model Selection ####
```{r VAR Short Term Model Building}
VARselect(ws.train.df, lag.max = 10, type= "both", season = 24, exogen = NULL) # AIC picked p = 6
ws.var = VAR(cbind(wsTrain.ap, wsTrain.ws), season = 24, type = "both", p = 6)
```

#### VAR Model Forecast ####
```{r VAR Short Term Forecast, results='hide'}
ws.var.fore = predict(ws.var, n.ahead = 24)
```

#### VAR Model Metrics ####
```{r VAR Short Term Metrics}
ws.var.mse = mean((wsTest$ActivePower - ws.var.fore$fcst$wsTrain.ap[1:24,1]) ^2)
paste0("VAR Short Term Predictions ASE: ", ws.var.mse)
```

#### VAR Model Plots ####
```{r VAR Short Term Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "24 Hour Power Forecast")
points(seq(277,300,1), ws.var.fore$fcst$wsTrain.ap[1:24,1], type = "l", col = "red")
```

```{r VAR Confidence Intervals}
fanchart(ws.var.fore, colors = brewer.pal(n=8, name="Reds"))
```

### Long Term Predictions ###

#### VAR Model Setup ####
```{r VAR Long Term Model Setup}
wlTrain.ap = wlTrain$ActivePower
wlTrain.ws = wlTrain$WindSpeed
wl.train.df = c(wlTrain.ap, wlTrain.ws)
```

#### VAR Model Selection ####
```{r VAR Long Term Model Building}
VARselect(wl.train.df, lag.max = 10, type= "both", season = 24, exogen = NULL) # AIC picked p = 6
wl.var = VAR(cbind(wlTrain.ap, wlTrain.ws), season = 24, type = "both", p = 6)
```

#### VAR Model Forecast ####
```{r VAR Long Term Forecast, results='hide'}
wl.var.fore = predict(wl.var, n.ahead = 72)
```

#### VAR Model Metrics ####
```{r VAR Long Term Metrics}
wl.var.mse = mean((wlTest$ActivePower - wl.var.fore$fcst$wlTrain.ap[1:72,1]) ^2)
paste0("VAR Long Term Predictions ASE: ", wl.var.mse)
```

#### VAR Model Plots ####
```{r VAR Long Term Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "72 Hour Power Forecast")
points(seq(229,300,1),wl.var.fore$fcst$wlTrain.ap[1:72,1], type = "l", col = "red")
```

##########################################
## Active Power Neural Network Analysis ##
##########################################

### Small Term Predictions ###

#### NN Model Setup ####
```{r NN Short Term Model Setup}
ws.train.df = data.frame(windspeed = ts(wsTrain$WindSpeed))
```

#### NN Model Building ####
```{r NN Short Term Model Building}
ws.train.fore.df = data.frame(part = ts(c(wsTrain$WindSpeed, ws.ws.mlp.fore$mean)))
ws.mlp.fit = mlp(ts(wsTrain$ActivePower, frequency = 24), reps = 50, comb = "median", xreg = ws.train.df) # 5 hidden node is best using rmse
ws.mlp.fit
```
```{r NN Short Term Model Structure}
plot(ws.mlp.fit)
```

#### NN Model Forecast ####
```{r NN Short Term Forecast, results='hide'}
ws.mlp.fore = forecast(ws.mlp.fit, h = 24, xreg = ws.train.fore.df)
```

#### NN Model Metrics ####
```{r NN Short Term Metrics}
ws.mlp.mse = mean((wsTest$ActivePower - ws.mlp.fore$mean) ^2)
paste0("Neural Network Short Term Predictions ASE: ", ws.mlp.mse)
```

#### NN Model Plots ####
```{r NN Short Term Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "24 Hour Power Forecast")
points(seq(277,300,1), ws.mlp.fore$mean, type = "l", col = "red")
```

### Long Term Predictions ###

#### NN Model Setup ####
```{r NN Long Term Model Setup}
wl.train.df = data.frame(windspeed = ts(wlTrain$WindSpeed))
```

#### NN Model Building ####
```{r NN Long Term Model Building}
wl.train.fore.df = data.frame(part = ts(c(wlTrain$WindSpeed, wl.ws.mlp.fore$mean)))
wl.mlp.fit = mlp(ts(wlTrain$ActivePower, frequency = 24), reps = 50, hd=4, comb = "median",xreg = wl.train.df) # 4 hidden node is best using rmse
wl.mlp.fit
```
```{r NN Long Term Model Structure}
plot(wl.mlp.fit)
```

#### NN Model Forecast ####
```{r NN Long Term Forecast, results='hide'}
wl.mlp.fore = forecast(wl.mlp.fit, h = 72, xreg = wl.train.fore.df)
```

#### NN Model Metrics ####
```{r NN Long Term Metrics}
wl.mlp.mse = mean((wlTest$ActivePower - wl.mlp.fore$mean) ^2)
paste0("Neural Network Long Term Predictions ASE: ", wl.mlp.mse)
```

#### NN Model Plots ####
```{r NN Long Term Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "72 Hour Power Forecast")
points(seq(229,300,1), wl.mlp.fore$mean, type = "l", col = "red")
```

####################################
## Active Power Ensemble Analysis ##
####################################

### Small Term Predictions ###

#### Ensemble Model Forecast ####
```{r Ensemble Short Term Forecast, results='hide'}
ws.ens.fore  = (ws.var.fore$fcst$wsTrain.ap[,1] + ws.mlp.fore$mean)/2
```

#### Ensemble Model Metrics ####
```{r Ensemble Short Term Forecast Metrics}
ws.ens.mse = mean((wsTest$ActivePower - ws.ens.fore) ^2)
paste0("Ensemble Short Term Predictions ASE: ", ws.ens.mse)
```

#### Ensemble Model Plots ####
```{r Ensemble Short Term Forecast Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "24 Hour Power Forecast")
points(seq(277,300,1), ws.ens.fore, type = "l", col = "red")
```

#### Long Term Predictions ####

#### Ensemble Model Forecast ####
```{r Ensemble Long Term Forecast, results='hide'}
wl.ens.fore  = (wl.var.fore$fcst$wlTrain.ap[,1] + wl.mlp.fore$mean)/2
```

#### Ensemble Model Metrics ####
```{r Ensemble Long Term Metrics}
wl.ens.mse = mean((wlTest$ActivePower - wl.ens.fore) ^2)
paste0("Ensemble Long Term Predictions ASE: ", wl.ens.mse)
```

#### Ensemble Model Plots ####
```{r Ensemble Long Term Plots}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,300), ylab = "Power", main = "72 Hour Power Forecast")
points(seq(229,300,1), wl.ens.fore, type = "l", col = "red")
```

####################
## Final Forecast ##
####################

#### ARMA(0,0,0) s=12 Model Forecast ####
```{r ARMA(0,0,0) Final Forecast, results='hide', fig.show = 'hide'}
final.arimaOne.fore = fore.arima.wge(wind$ActivePower, s = 24, n.ahead = 72)
```

#### ARMA(1,0,0) s=12 Model Forecast ####
```{r ARMA(1,0,0) Final Forecast, results='hide', fig.show = 'hide'}
final.s24 = artrans.wge(wind$ActivePower, c(rep(0,23), 1))
final.s24.aic = aic.wge(final.s24, type = 'bic')
final.arimaTwo.fore = fore.arima.wge(wind$ActivePower, phi = final.s24.aic$p, s = 24, n.ahead = 72)
```

#### VAR Model Forecast ####
```{r VAR Final Forecast, results='hide'}
final.ap = wind$ActivePower
final.ws = wind$WindSpeed
final.train.df = c(final.ap, final.ws)
final.var = VAR(cbind(final.ap, final.ws), season = 24, type = "both", p = 6)
final.var.fore = predict(final.var, n.ahead = 72)
```

#### Wind Speed Forecast ####
```{r Wind Speed Final Forecast, results='hide'}
final.ws.mlp.fit = mlp(ts(wind$WindSpeed, frequency = 24), reps = 50, comb = "median")
final.ws.mlp.fore = forecast(final.ws.mlp.fit, h = 72)
```

#### MLP Model Building and Forecast ####
```{r NN Final Forecast, results='hide'}
final.train.df = data.frame(windspeed = ts(wind$WindSpeed))
final.train.fore.df = data.frame(part = ts(c(wind$WindSpeed, final.ws.mlp.fore$mean)))
final.mlp.fit = mlp(ts(wind$ActivePower, frequency = 24), reps = 50, hd=4, comb = "median", xreg = final.train.df)
final.mlp.fore = forecast(final.mlp.fit, h = 72, xreg = final.train.fore.df)
```

#### Ensemble Model Forecast ####
```{r Ensemble Final Forecast, results='hide'}
final.ens.fore  = (final.var.fore$fcst$final.ap[,1] + final.mlp.fore$mean)/2
```

#### Final Plot ####
```{r Final Plot}
plot(seq(1,300,1), wind$ActivePower, type = "l",xlim = c(0,372), ylab = "Power", xlab = "Time", main = "72 Hour Power Forecast")
points(seq(301,372,1), final.arimaTwo.fore$f, type = "l", col = "blue", lwd=1.5)
points(seq(301,372,1), final.var.fore$fcst$final.ap[1:72,1], type = "l", col = "green", lwd=1.5)
points(seq(301,372,1), final.mlp.fore$mean, type = "l", col = "purple", lwd=1.5)
points(seq(301,372,1), final.ens.fore, type = "l", col = "orange", lwd=1.5)
points(seq(301,372,1), final.arimaOne.fore$f, type = "l", col = "red", lwd=2)
```